### 2016-05-28 thesis/methods/methods.tex
tmp commit

\gls{rnaseq} data for the eight cancer types with sample \gls{bmi} information available were downloaded from the \gls{icgc} database.
The \gls{rnaseq} data for each of the eight cancer types were imported into \gls{r}.

@@

\section{Data processing -- \citet{Creighton2012} data}
\label{sec:dataprocessingcr}

@@

The raw microarray gene expression data and clinical data files of all *** samples from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
Raw data were imported into \gls{r}/RStudio for subsequent analyses.

@@

\subsection{Normalisation of raw \citet{Creighton2012} data}
\label{subsec:normcrdata}

@@

799 obesity associated gene probes identified by \citet{Creighton2012} were downloaded from (supplementary data??).
The list of gene probes were imported into \gls{r}.
The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Any duplicates and any gene symbols that were not present in both the Creighton data and the \gls{icgc} data were removed, which left *** unique obesity associated genes.

@@

The results presented by \citet{Creighton2012} were validated using the same raw data they have used in their study.
See  \cref{sec:dataprocessingcr} for details on how the data was processed before the analyses.

@@

\gls{svd} was applied to this data to obtain the decomposed matrices of the data.
%TODO: how do I explain how I got the \gls{metagene} scores?
The \gls{metagene} scores of the samples were obtained by taking the first principal component of the (eigenvector...?).
The \gls{metagene} scores were ranked based on the number of samples in the data (103 samples in Creighton data), and the values of the scores were between 0 and 1.

@@

\Gls{metagene} scores from \cref{subsec:icgctranscr} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap, for each cancer type.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

@@

%TODO: Not sure if I should cut this section to "see previous section for details"
\section{Data processing -- \citet{Fuentes-Mattei2014} data}
\label{sec:dataprocessingfm}

@@

The raw microarray gene expression data files of all *** samples from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
Raw data were imported into \gls{r}/RStudio for subsequent analyses.

@@

\subsection{Normalisation of raw \citet{Fuentes-Mattei2014} data}
\label{subsec:normfmdata}

The raw microarray data for all *** samples from the study were normalised using \gls{rma} method from the \textit{affy} package in \gls{r}.
The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.
This ensured that all of the genes were on the same scale as one another, allowing for comparison.

@@

The results presented by \citet{Fuentes-Mattei2014} were validated using the \citet{Creighton2012} data, as the clinical data of the samples were not available.
See \cref{subsec:rawdatacr,subsec:normcrdata} for details on how the data was processed before the analyses.

@@

%TODO: how do I explain how I got the \gls{metagene} scores?
The \gls{metagene} scores of the samples were obtained by taking the first principal component of the transformed matrix.
The \gls{metagene} scores were ranked based on the number of samples in the data (*** samples in \citet{Creighton2012} data), and the scores ranged from 0 to 1.

@@

The same \gls{metagene} scores from the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
Boxplot was plotted to show the association of the \gls{metagene} scores with the sample \gls{bmi} status, and scatter plot was plotted to show the association with the sample \gls{bmi} value.

@@

@@

All 103 samples from the \citet{Creighton2012} study were split into two groups (obese and non-obese) based on the sample \gls{bmi} status as defined by the \gls{who} classification (see \cref{tab:whobmiclass}).
The difference in the gene expression between these two sample groups were analysed by fitting a linear model to the data, using the \textit{limma} package in \gls{r}.
The top 799 differentially expressed gene probes that had p-value \textless{} 0.01 were identified.

>>>>

\gls{rnaseq} data for the eight cancer types with sample \gls{bmi} information available were downloaded from the \gls{icgc} database.
The \gls{rnaseq} data for each of the eight cancer types were imported into \gls{r}.
%TODO: add in how I processed the ICGC data format into genes by samples.

@@

\section{Data processing -- Creighton data}
\label{sec:dataprocessingcr}

@@

The raw microarray gene expression data files of all *** samples from the \citet{Creighton2012} study were downloaded from the \gls{geo} database (\gls{geo} accession ID: GSE24185).
Clinical data of the samples were obtained from the supplementary table 1 from their paper.
Raw data were imported into \gls{r}/RStudio for subsequent analyses.

@@

\subsection{Normalisation of raw Creighton data}
\label{subsec:normcrdata}

@@

799 obesity associated gene probes identified by \citet{Creighton2012} were obtained from the supplementary data file 1 from their paper.
The list of gene probes were imported into \gls{r}.
The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Any duplicates and any gene symbols that were not present in both the Creighton data and the \gls{icgc} data were removed, which left *** unique obesity associated genes.

@@

The results presented by \citet{Creighton2012} were validated using the same raw data they have used in their study.
See  \cref{sec:dataprocessingcr} for details on how the raw data was accessed and processed.

@@

\gls{svd} was applied to this data to obtain the decomposed matrices of the data.
%TODO: how do I explain how I got the \gls{metagene} scores?
The \gls{metagene} scores of the samples were obtained by taking the first eigenvector of the (orthogonal matrix...?).
The \gls{metagene} scores were ranked based on the number of samples in the data (103 samples in Creighton data), and the values of the scores were between 0 and 1.

@@

\Gls{metagene} scores from \cref{subsec:icgctranscr} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap for each cancer type.
The data from \cref{subsec:datprocicgc} were used for the heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

@@

%TODO: Not sure if I should cut this section to "see previous section for details"
\section{Data processing -- Fuentes-Mattei data}
\label{sec:dataprocessingfm}

@@

The raw microarray gene expression data files of all *** samples from the  \citet{Fuentes-Mattei2014} study were downloaded from the \gls{geo} database (include website name/GEO ID here).
Clinical data for the samples in this study was not available.
Raw data were imported into \gls{r}/RStudio for subsequent analyses.

@@

\subsection{Normalisation of raw Fuentes-Mattei data}
\label{subsec:normfmdata}

The raw microarray data for all *** samples from the study were normalised using \gls{rma} method from the \textit{affy} package in \gls{r}.
The data was then standardised so that each of the genes had a mean ($\mu$) of 0 and a standard deviation ($\sigma$) of 1.

@@

The results presented by \citet{Fuentes-Mattei2014} were validated using the Creighton data, as the clinical data of the samples were not available.
See \cref{subsec:rawdatacr,subsec:normcrdata} for details on how the data was processed.

@@

%TODO: how do I explain how I got the \gls{metagene} scores?
The \gls{metagene} scores of the samples were obtained by taking the first eigenvector of the transformed Creighton data.
The \gls{metagene} scores were ranked based on the number of samples in the data (*** samples in Creighton data), and the scores ranged from 0 to 1.

@@

The same \gls{metagene} scores from the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values.
Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.

@@

\subsection{\gls{icgc} data processing}
\label{subsec:datprocicgcfm}

All of the \gls{icgc} data were processed in a similar fashion as in \cref{subsec:datprocicgc}, but the obesity associated genes from \cref{subsec:fmobsgene} were used instead.

\subsection{Transformation matrix}
\label{subsec:icgctransmatfm}

The same transformation matrix from \cref{subsec:transmatfm} was used.

\subsection{Transformed \gls{icgc} data and metagene}
\label{subsec:icgctransfm}

Each of the eight cancer data from \cref{subsec:datprocicgcfm} were transformed using the transformation matrix from \cref{subsec:transmatfm}.
Metagene scores were ranked as in \cref{subsec:icgctranscr}.

\subsection{Fuentes-Mattei metagene and sample gene expression in \gls{icgc} data}
\label{subsec:fmmetaicgcge}

\Gls{metagene} scores from \cref{subsec:icgctransfm} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap for each cancer type.
The data from \cref{subsec:datprocicgcfm} were used for the heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{Creighton metagene and sample \gls{bmi} in \gls{icgc} data}
\label{subsec:fmmetaicgcge}

The same \gls{metagene} scores as in the previous section were used to compare the \gls{metagene} scores with the samples' \gls{bmi} statuses and values for each cancer type.
Box and scatter plots, and any statistical values associated with these plots were calculated as in \cref{subsec:crmetaicgcge}.

@@

\subsection{Data processing and subsetting}
\label{subsec:subdatadeg}

Raw data from \cref{subsec:rawdatacr,subsec:normcrdata} were used in the analysis.
Together with the raw data, few other data sets were created based on the raw data: residual data, Caucasian-only raw data, and Caucasian-only residual data.

Residual data was created from the raw data after it has been controlled for all clinical variables, except those related to \gls{bmi}.
These clinical variables included age, sex, ethinicity, tumour grade, \gls{pr}, \gls{er}, \gls{her2} statuses, and \gls{ln} status.

Caucasian-only data was created by taking only the Caucasian samples from the raw data, which reduced the sample size from 103 samples to 77 samples.
Caucasian-only residual data was created using the same approach as with the residual data: the data that remained after all the non-\gls{bmi} clinical variables were controlled in the Caucasian-only raw data.

\subsection{Group classification}
\label{subsec:crdegclass}

In each of the datasets from \cref{subsec:subdatadeg}, all of the samples were split into two groups: obese and non-obese groups.
These two groups were classified based on the sample \gls{bmi}, where the obese groups were defined as samples with \gls{bmi}  \textgreater{} 30, and non-obese group were defined as samples with \gls{bmi}  \textless{} 30 (see \cref{tab:whobmiclass}).

\subsection{Gene expression analysis}
\label{subsec:crdeg}

The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Where there was a duplicated gene in the data, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.

The difference in the gene expression between these two sample groups were analysed by fitting a linear model to the data, using the \textit{limma} package in \gls{r}.
The top 799 differentially expressed gene probes that had p-value \textless{} 0.01 were identified.

\newpage
